---
title: "wrangle_turnout"
output: html_document
---

Wrangling election and precinct variables in December data (use dec_wrangled.Rda file)

```{r}

load("/Users/eliane/Downloads/dec_wrangled (1).Rda")

overall_turnout_dec <- overall_turnout
overall_turnout_dec$election <- str_remove(overall_turnout_dec$election, "voted_in_") 
overall_turnout_dec$election <- str_remove(overall_turnout_dec$election, "_voting_method")
overall_turnout_dec$precinct_name <- str_remove(overall_turnout_dec$precinct_name, "BALTIMORE CITY - ")

overall_turnout_dec <- separate(overall_turnout_dec, col = election, 
         into=c('election_year','election_type1',
                'election_type2'), sep = '_')

overall_turnout_dec2 <- unite(overall_turnout_dec, 
                              col = "election",
                              election_type1, election_type2,
                                sep = " ")

#unique(overall_turnout_dec2$election)
```

Wrangling election variables in May 2020 data 

```{r}
may_2020 <- read_csv("/Users/eliane/Desktop/R Folder/Data Hackathon/overall_turnout_may.csv")

may_2020$election <- tolower(may_2020$election)

may_2020 <- separate(may_2020, col = election, 
                     into=c('election_type1',
                            'election_type2', 'election_year'), sep = '_')

may_2020_2 <- unite(may_2020, col = "election",
                              election_type1, election_type2,
                                sep = " ")

# unique(may_2020$election_type2)
# unique(overall_turnout_dec$election_type2)
```

After cleaning up the election and precinct variables individually, 
ensure that the precincts match up between datasets

First chunk, ensure that there are no precincts that don't exist in the other dataset
Second chunk, pull the "unique" precinct values for each dataset and ensure they're the same
Third chunk, find precinct values that do not exist in the other

```{r}
MayPrecinct <- may_2020$Precinct
DecPrecinct <- overall_turnout_dec$precinct_name
MayPrecinct[!(MayPrecinct %in% DecPrecinct)]
DecPrecinct[!(DecPrecinct %in% MayPrecinct)]

head(MayPrecinct)
head(DecPrecinct)
May_pre <-sort(unique(MayPrecinct))
Dec_pre <-sort(unique(DecPrecinct))
identical(May_pre, Dec_pre)

setdiff(MayPrecinct, DecPrecinct)
# make May and Dec match
         
```

Final steps - bind the cleaned up dec and may data 
The election variables are now compatible between datasets

```{r}

names(overall_turnout_dec2)
names(may_2020_2)

overall_turnout_dec2 <- overall_turnout_dec2 %>% 
  select(election_year, election, "precinct" = precinct_name,
         "voted_no" = No_vote, "voted_yes" = Total_vote,
         turnout_perc)

may_2020_2 <- may_2020_2 %>% select(election_year, election,
                                    "precinct" = Precinct, voted_no, voted_yes,
                                    turnout_perc)

bind_overall_turnout <- bind_rows(may_2020_2, overall_turnout_dec2, .id= "source")
bind_overall_turnout <- bind_overall_turnout %>% 
  mutate(source = case_when(source == 1  ~ "may",
                            source == 2 ~ "dec"))

#View(bind_overall_turnout %>% filter(election_year == 2016) %>% filter(precinct == 25003))
```
