---
title: "Voter Data Import"
output: 
  html_document:
    toc: TRUE
    toc_float: true
---


# General Thoughts

To use googlesheets4:
1) Data needs to be resaved as googlesheet not xlsx 
2) Need at least commenting permission to access

I am not sure what is the most sustainable option for gathering new data... maybe just having a computer powerful enough to download the data? is dta the raw format? Should ask about this.

XLSX version had messed up birth dates - so that had to be fixed.


```{r, eval = FALSE}
library(googledrive)
library(googlesheets4)
library(tidyverse)
```

# Importing voter Data from 2020 

## May data

First trying with xlsx:

```{r,  eval = FALSE}
#Data from May imported by Carrie on her machine - not including data on GitHub for security reasons
Maypart1<- readxl::read_xlsx(here::here("data", "voter_data", "may_2020", "May_part1.xlsx"))
#This caused errors with birthday thus I saved it as a csv in excel using the xlsx file which is used in the next chunk
```

Next using csv for part 1:

```{r, eval = FALSE}
# import may part 1 (saved as csv so birthday is fixed) 
# specify character for nonstandard address col
Maypart1<- read_csv(here::here("data", "voter_data", "may_2020", "May_part1.csv"), 
                    col_types = cols(NonStandardAddress = col_character()))

# check problems
if(nrow(problems(Maypar12)) == 0) print("ALL Good!") else(("Check data"))
#all good!
```

Need to not read first row as a header for part 2, instead we will use the header of part 1 as the header so that the first row's data is included.

```{r, eval = FALSE}
# import may part 2 (saved as csv so birthday is fixed) 
# specify character for nonstandard address col
Maypart2<- read_csv(here::here("data", "voter_data", "may_2020", "May_part2.csv"), col_names = colnames(Maypart1),
                    col_types = cols(NonStandardAddress = col_character()))

# check problems
problems(Maypart2)

if(nrow(problems(Maypart2)) == 0) print("ALL Good!") else(("Check data"))
#all good!
```



If you wanted to use googlesheets4 could do this... but security risk to have URL out in public. We want this data to only be used the way we intend.

```{r, eval = FALSE}
# removing URL for security reasons
Maypart2_url <- "{URL of link goes here}"
Maypart2 <-googlesheets4::read_sheet(Maypart2_url)
```

Put two may data peices together and save:

```{r, eval = FALSE}
May <-bind_rows(Maypart1, Maypart2)
save(May, file = here::here("data", "voter_data","may_2020",  "may_2020_voter_data.rda"))
set.seed(1234)
May_sample <-May %>% slice_sample(n = 50)
save(May_sample, file = here::here("data", "voter_data","may_2020",  "may_2020_voter_data_random_subset.rda"))
```



## December data

```{r, eval = FALSE}
Dec_2020 <-read_csv(here::here("data", "voter_data", "dec_2020", "Dec_2020.csv"))
problems(Dec_2020) %>%pull(expected) %>% unique()
#Looks like some people dont have the day for their birthday, just year and month and 00 for day.
#issue with true false expectation
colnames(Dec_2020)[problems(Dec_2020) %>%filter(expected !="date in ISO8601" ) %>% pull(col) %>% unique()]
# only appears to be issue with voting method or party -so should be OK

save(Dec_2020, file = here::here("data", "voter_data","dec_2020",  "dec_2020_voter_data.rda"))

set.seed(1234)
Dec_sample <-Dec_2020 %>% slice_sample(n = 50)
save(Dec_sample, file = here::here("data", "voter_data","dec_2020",  "dec_2020_voter_data_random_subset.rda"))
```

# Instructions to load data for wrangling
Note this is set up for having the data in a data directory:

- that has a subdirectory called voter_data
- that has a subdirectory called may_2020 and dec_2020


To get less data to practice with use this code:

```{r}
load(file = here::here("data", "voter_data","may_2020",  "may_2020_voter_data_random_subset.rda"))
load(file = here::here("data", "voter_data","dec_2020",  "dec_2020_voter_data_random_subset.rda"))

```


***Looks like some people dont have the day for their birthday, just year and month and 00 for day - something to keep in mind.***


To load all of the May & Dec 2020 data:

```{r}
load(file = here::here("data", "voter_data","may_2020",  "may_2020_voter_data.rda"))
load(file = here::here("data", "voter_data","dec_2020",  "dec_2020_voter_data.rda"))
```




